Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-22 WSSSPE3/report/WSSSPE3_review.tex
edits in new text

Returning to the topic of his talk, Turk invited the audience to picture
scientific software as a flower on a landscape under the Sun, which may
represent in this analogy a number of measurable factors such as number of
citations; growth of a community and number of contributors; amount of funding;
prestigious prizes awarded; stability of the community in terms of leadership
transitions, serving community needs, not breaking test suites, and performance
on new architectures. But all these metrics are strictly speaking \emph{proxies}
for the values and the impact scientific software bears. What we can measure
does not give us direct insight---it just gives us proxies of insight.

Turk then moved onto various different definitions of sustainability. His most
favourite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding on new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software.
Some funder Turk heard talked about sustainability as getting funded. Turk also
recalled that Greg Wilson said in WSSSPE 1.1 in his view sustainable software
was software that continued to give the same results over time. A last measure
of sustainability Turk presented was the ability to transition between different people
developing and using a piece of software.

>>>>

Returning to the topic of his talk, Turk invited the audience to picture
scientific software as a flower on a landscape under the Sun, which may
represent a number of measurable factors such as number of
citations; growth of a community and number of contributors; amount of funding;
prestigious prizes awarded; stability of the community in terms of leadership
transitions, serving community needs, not breaking test suites, and performance
on new architectures. But all these metrics are strictly speaking \emph{proxies}
for the values and the impact scientific software bears. What we can measure
does not give us direct insight---it just gives us proxies of insight.

Turk then moved onto various different definitions of sustainability. His 
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software.
A funder Turk heard talked about sustainability as continuing to get funded. Turk also
recalled that Greg Wilson said in WSSSPE 1.1 that his view of sustainable software
was software that continued to give the same results over time. A last measure
of sustainability Turk presented was the ability to transition between different people
developing and using a piece of software.
